\documentclass[conference,nofonttune]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}
% Times-compatible text and math fonts without missing TU/ptm shapes
\usepackage{newtxtext,newtxmath}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{{Title}
    \thanks{Term Paper submitted for SC6101 Introduction to Blockchain, NTU.}
}

\author{
    \IEEEauthorblockN{Chen Sibei}
    \IEEEauthorblockA{G2509013K \\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        SCHEN050@e.ntu.edu.sg}\\
    \IEEEauthorblockN{Shan Xingzhong}
    \IEEEauthorblockA{G2509363F\\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        shan0159@e.ntu.edu.sg}

    \and
    \IEEEauthorblockN{Feng Jiahao}
    \IEEEauthorblockA{G2509193G \\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        FENG0207@e.ntu.edu.sg}\\
    \IEEEauthorblockN{Tian Zhenman}
    \IEEEauthorblockA{G2509210A \\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        ZHENMAN001@e.ntu.edu.sg}

    \and
    \IEEEauthorblockN{Mo Fan}
    \IEEEauthorblockA{Student's Matriculation Number \\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        Email address of Student}\\
    \IEEEauthorblockN{Zhang Boyu}
    \IEEEauthorblockA{G2509012B \\
        \textit{Graduate College}\\
        \textit{Nanyang Technological University}\\
        Singapore \\
        BZHANG037@e.ntu.edu.sg}
}

\maketitle

\begin{abstract}
    As stable units of account and mediums of exchange, stablecoins are widely used in decentralized finance (DeFi). However, the processes of stablecoin trading and collateral management typically inherit the inherent transparency of public blockchains, posing significant privacy leakage to both individual users and institutional participants. Given the widespread use of stablecoins in payment systems, payroll settlements, merchant settlements, over-the-counter (OTC) trading, and cross-border transfers, this paper considers privacy as a primary concern in stablecoin applications. In these areas, transaction correlation and identity inference can directly expose sensitive financial information. We examine key privacy challenges in the stablecoin ecosystem, including on-chain traceability and address correlation, privacy vulnerabilities arising from collateral and liquidation mechanisms, cross-layer and cross-chain data correlation, and regulatory compliance requirements (e.g., Know Your Customer (KYC)/Anti-Money Laundering (AML)) that may conflict with user privacy expectations. We analyze relevant threat models and real-world impacts and propose comprehensive mitigation strategies that include cryptographic solutions (e.g., zero-knowledge proofs), protocol-level design improvements, wallet and user experience enhancements, and privacy-preserving compliance architectures tailored to each challenge.
\end{abstract}

\begin{IEEEkeywords}
    Stablecoins, Blockchain Privacy, DeFi, Privacy-Preserving Techniques, Zero-Knowledge Proofs, KYC/AML Compliance, Address Linkability, Transaction Traceability, Collateralized Stablecoins, Cross-Chain Privacy, Layer 2 Scaling, MEV, Regulatory Compliance, Virtual Asset Service Providers, Cryptographic Protocols
\end{IEEEkeywords}

\section{Introduction}
Stablecoins have evolved from a niche payment tool to a core settlement layer for digital asset markets. Their fiat-pegged design makes them a preferred unit of account for trading, remittances, and treasury operations. Recent policy and market commentary suggests that stablecoins mediate a large share of crypto trading and settlement activity \cite{waller2025,theblock2025}. Still, public verifiability, composability with smart contracts, and frictionless transferability - these very properties that make stablecoins scalable and interoperable also create a privacy landscape that is often worse than that of native cryptocurrencies.

Privacy challenges in stablecoins differ from those in Bitcoin-like systems in both technology and governance. On the technical side, most major stablecoins run in account-based environments on Ethereum-compatible networks. In practice, address reuse is widespread, and ERC-20 transfer events expose fine-grained metadata through on-chain logs. These features enable powerful transaction-graph analysis and clustering and can be further strengthened by network-layer/Remote Procedure Call (RPC) side channels.
On the institutional side, stablecoins often add actors with stronger surveillance and control capabilities than those found in permissionless cryptocurrencies. Centralized issuers and regulated intermediaries typically hold off-chain identity records collected through KYC, and many stablecoin contracts support administrative actions such as freezing or blacklisting. As a result, pseudonymity can be fragile even when users follow good on-chain hygiene. At the same time, stablecoin adoption is increasingly shaped by regulatory and compliance requirements—such as KYC/AML controls and sanctions screening—which can compel identity disclosure and widen the overall surveillance surface. These requirements are shaped by frameworks including the Financial Crimes Enforcement Network (FinCEN), the Financial Action Task Force (FATF), the Office of Foreign Assets Control (OFAC), and the Markets in Crypto-Assets Regulation (MiCAR) \cite{finCEN2019,fatf2021,ofac2021,micar2023}.

This paper provides a structured analysis of stablecoin privacy leakage, focusing on the interplay between transparency and compliance requirements. We divide our discussion into four dimensions frequently encountered in real-world deployments and empirical studies. First, we examine on-chain traceability and address linkability, including clustering heuristics that map aliases to entities and side-channel attacks. Then, we analyze privacy leakage arising from collateralized stablecoin architectures, where vault status, liquidation thresholds, and auction mechanisms expose sensitive financial information that can be maliciously exploited. Third, we discuss cross-layer and cross-chain connections between L1/L2 layers and bridges, where bridging events and time/amount heuristics act as linkable anchors, connecting user activities across different domains. Fourth, we examine compliance-driven identity disclosure, using fiat-backed stablecoins as an example to demonstrate how regulatory changes and issuer privileges erode privacy over time.

\paragraph{Terminology and scope}
In this paper, we use \emph{privacy leakage} as an umbrella term that covers (i) \emph{linkability} (linking addresses, users, and activities) and (ii) \emph{confidentiality loss} (exposure of amounts, positions, or relationships). We describe adversaries using a common threat-model vocabulary: a Global Passive Adversary (GPA), a Network-Level Adversary (NLA), and a privileged compliance actor (e.g., issuer/Virtual Asset Service Provider (VASP)). We also group impacts into four recurring dimensions: \emph{identity exposure}, \emph{financial exposure} (balances or positions), \emph{strategic exploitation} (e.g., Mining Efficiency Value (MEV), liquidation sniping), and \emph{compliance/control risk} (freezing, blacklisting, censorship).

Our contributions are threefold. (1) We define realistic threat models tailored to stablecoins, including both conventional global observers and privileged issuer/compliance adversaries. (2) We summarize representative attack and inference mechanisms across transfers, positions, and cross-domain movements, and highlight their privacy impacts on both retail and institutional users. (3) We survey mitigation directions that aim to balance privacy with accountability, including stealth address schemes, privacy pools with selective disclosure, and minimal-disclosure cryptographic proofs that preserve verifiable constraints while reducing data exposure \cite{baldimtsi2024,duffie2025}.

The remainder of this paper is organized as follows. Section~II analyzes on-chain traceability and address linkability. Section~III discusses privacy leakage from collateral, liquidation, and position transparency. Section~IV examines cross-layer and cross-chain correlation for L2s and bridges. Section~V focuses on compliance-driven identity disclosure and design trade-offs. Section~VI concludes with open challenges and directions for privacy-preserving yet regulation-compatible stablecoin architectures.

Table~\ref{tab:overview} provides an overview of the threat surfaces, typical impacts, and mitigation directions discussed in the following sections.

\begin{table*}[t]
    \centering
    \small
    \setlength{\tabcolsep}{4pt}
    \renewcommand{\arraystretch}{1.15}
    \caption{Summary of stablecoin privacy risks, impacts, and mitigations}
    \label{tab:overview}
    \begin{tabular}{p{2.4cm} p{4.7cm} p{4.7cm} p{3.8cm}}
        \hline
        \textbf{Risk class}                                              & \textbf{Main leakage channel} & \textbf{Typical impact} & \textbf{Mitigation direction} \\
        \hline
        On-chain traceability                                            &
        ERC-20 logs, address reuse, clustering, RPC timing               &
        Identity linkage, payment routine profiling, MEV targeting       &
        Stealth addresses, privacy pools, RPC hygiene                                                                                                              \\
        \hline
        Collateral \& liquidation transparency                           &
        Public vault states, thresholds, auctions, mempool observability &
        Position exposure, liquidation sniping, adverse execution        &
        Minimal-disclosure proofs, auction design, commit--reveal                                                                                                  \\
        \hline
        Cross-layer/cross-chain correlation                              &
        Bridge events as anchors, timing/amount matching                 &
        End-to-end path reconstruction, business relationship exposure   &
        Batching/delays, metadata minimization, cross-domain privacy                                                                                               \\
        \hline
        Compliance-driven disclosure                                     &
        KYC datasets, blacklist/freeze controls, sanctions screening     &
        De-pseudonymization, censorship risk, selective freezing         &
        Selective disclosure, verifiable credentials, compliant privacy                                                                                            \\
        \hline
    \end{tabular}
\end{table*}

\section{On-Chain Traceability and Address Linkability}
Stablecoins on Ethereum and its compatible networks—primarily ERC-20 tokens such as USD Coin (USDC), Tether (USDT), and DAI—have become the cornerstone of the digital asset ecosystem. However, this convenience is built upon an extremely transparent architecture—the account-based model. Unlike Bitcoin's Unspent Transaction Output (UTXO) model, the account-based model is designed to encourage address reuse and lacks a native change mechanism. This architectural characteristic, combined with the event logs mechanism of the ERC-20 standard, subjects stablecoin transactions to more severe privacy challenges than native cryptocurrencies.

\subsection{Threat Model}
When analyzing stablecoin privacy, it is essential to first define the scope and perspective of an attacker's capabilities. The academic community typically categorizes adversaries targeting blockchain privacy into three tiers: global passive adversaries, network-layer adversaries, and centralized issuers/regulatory adversaries.

\paragraph{Global Passive Adversary (GPA)}
GPA is assumed to possess a complete copy of the blockchain ledger (i.e., running an archival node), capable of parsing all block data, transaction records, and event logs generated by smart contracts since the genesis block.
\begin{itemize}
    \item \textbf{Capability:} The GPA constructs a complete transaction graph $G(V, E)$, where $V$ represents addresses and $E$ represents token or ETH transfers.
    \item \textbf{Goal:} To map multiple pseudonyms to a single real-world entity using clustering heuristics.
    \item \textbf{Reality:} This model corresponds to blockchain analytics firms (e.g., Chainalysis) \cite{chainalysis2025} and academic researchers. Victor (2020) demonstrated that GPAs can effectively cluster Ethereum users by exploiting deposit address reuse \cite{victor2020}.
\end{itemize}

\paragraph{Network-Level Adversary (NLA)}
Unlike GPA, which focuses solely on on-chain data, NLA monitors the underlying Peer-to-Peer (P2P) network or RPC communication traffic.
\begin{itemize}
    \item \textbf{Capability:} NLA attempts to associate transaction broadcast times or network packet characteristics with physical IP addresses.
    \item \textbf{Latest Threat:} Wang et al. (2025) proposed the "Time Tells All" attack model, demonstrating that NLA no longer requires operating supernodes or controlling large numbers of P2P connections. Instead, it can achieve de-anonymization simply by monitoring encrypted Transmission Control Protocol (TCP) traffic—such as through Internet Service Provider (ISP) or local network eavesdroppers \cite{wang2025}.
\end{itemize}

\paragraph{Privileged Issuer Adversary}
This is a threat model specific to stablecoins, particularly fiat-collateralized stablecoins.
\begin{itemize}
    \item \textbf{Capability:} Issuers (such as Circle and Tether) not only possess an on-chain view of GPA but also control off-chain KYC data. Furthermore, they hold god-like privileges at the smart contract level, including blacklisting and freezing addresses.
    \item \textbf{Impact:} This privilege enables taint analysis to be enforced. Once an address is flagged, other addresses that have interacted with it—even if the interaction occurred several hops prior—may face the risk of asset freezing.
\end{itemize}

\subsection{Attack and Inference Mechanisms}
The core of de-anonymization attacks lies in breaking the unlinkability between addresses. Under the Ethereum account model, academia has developed multiple high-precision heuristic algorithms. Combined with network-layer side-channel attacks, these techniques render stablecoin users' privacy virtually impossible to conceal.

\paragraph{Clustering Heuristics in the Account Model}
In Bitcoin, the Multi-input Heuristic serves as the gold standard for clustering. However, Ethereum transactions typically involve only a single sender, rendering Bitcoin's clustering methods ineffective. For Ethereum, researchers have proposed novel heuristic algorithms based on account behavior patterns.
\begin{itemize}
    \item \textbf{Deposit Address Reuse:}
          In his paper \textit{Address Clustering Heuristics for Ethereum}, presented at the premier conference Financial Cryptography in 2020, Friedhelm Victor systematically quantified for the first time the privacy leakage resulting from address reuse.
    \item \textbf{Gas Payer Linking (The ``Sugar Daddy'' Attack):}
          Identified by B\'{e}res et al. (2020), this is a potent attack against ERC-20 privacy \cite{beres2020}. Since stablecoins cannot pay for their own execution gas directly, a fresh address ($Address_{New}$) receiving USDC needs Ether (ETH) to move it. Users typically fund this gas from a main wallet ($Address_{Main}$) or a Centralized Exchange (CEX).
          $$ Address_{Main} \xrightarrow{\text{ETH}} Address_{New} $$
          This funding transaction creates a visible link. Even if the stablecoins received by $Address_{New}$ are from a private source, the gas funding trail reveals the entity relationship, creating a ``star topology'' in the transaction graph.
    \item \textbf{Behavioral Fingerprinting:}
          Reuter (2025) highlights that time-zone analysis of transaction timestamps can geo-locate users \cite{reuter2025}.
          Furthermore, the set of contracts a user approves (via the \texttt{approve()} function) creates a unique fingerprint.
          If two addresses approve the same specific sequence of DeFi protocols (e.g., Uniswap V3, Aave, and a niche Non-Fungible Token (NFT) market), they are likely controlled by the same user.
\end{itemize}

\paragraph{RPC/P2P Side-Channel Attacks}
Although on-chain clustering can only associate addresses with "entities," network-layer attacks can link addresses to "Internet Protocol (IP) addresses," thereby enabling physical geolocation.
\begin{itemize}
    \item \textbf{"Time Tells All" Attack:}
          Wang et al. (2025) presented a de-anonymization attack targeting RPC users in a USENIX Security/CCS-level research paper \cite{wang2025}.
    \item \textbf{"P2P Topology" Attacks:}
          Early studies focused on message propagation delays at the P2P layer.
          While the impact is mitigated for users relying on RPC, further research by Heimbach et al. (2025) demonstrated that by analyzing proof messages (Attestations) within TCP connections, the physical locations of over 15\% of Ethereum validator nodes could be identified \cite{heimbach2025}.
    \item \textbf{"Specific Leakage in ERC-20 Graphs":}
          The transfer patterns of stablecoins exhibit significant topological differences from native ETH patterns, providing entry points for specific analyses.
          Josenhans et al. (2025) noted in their analysis of ERC-20 scam token networks that the peel chain pattern—where large funds gradually diminish through successive transfers, akin to peeling an onion—is highly prevalent in token laundering \cite{josenhans2025}.
          Using graph embedding technology, researchers achieved 88.7\% accuracy in identifying anomalous token transfer networks.
          For stablecoins, this pattern typically indicates users attempting to obscure fund paths via `bridge' addresses.
\end{itemize}

\subsection{Impact}
The combined application of the aforementioned attack mechanisms has resulted in stablecoin users facing a multidimensional collapse of privacy:

\paragraph{Physical Identity Exposure}
By combining CEX KYC data (obtained through deposit address reuse) with ISP IP data (obtained via RPC side-channels), attackers can fully map on-chain pseudonyms to real-world legal entities.

\paragraph{Wealth and Behavior Insights}
Reuter (2025) research demonstrates how on-chain behavioral analysis can map stablecoin flows to specific geographic regions \cite{reuter2025}.
When this macro-level insight falls into the hands of criminals, it may enable targeted phishing attacks or even offline extortion against high-value holders.

\paragraph{MEV and Front-Running Transactions}
High-frequency traders leverage address clustering data to build `whale' profiles.
When flagged stablecoin holders initiate transfers, MEV bots exploit the gas bidding mechanism to execute front-running or sandwich attacks, causing direct financial losses to users.

\subsection{Mitigations}
To address these threats, the academic community and developer community have proposed various cryptography-based mitigation solutions, primarily focusing on stealth addresses and privacy pool technologies.

\paragraph{Stealth Addresses}
Stealth addresses are designed to sever on-chain links between senders and recipients, preventing third parties from identifying who received the funds.
\begin{itemize}
    \item \textbf{Mechanism Principle:}
          Based on the ERC-5564 standard and the Umbra protocol, stealth addresses utilize Elliptic Curve Diffie-Hellman (ECDH) key exchange. The process begins with the recipient publishing a "Meta-Address." The sender uses this Meta-Address to generate a temporary, one-time Stealth Address and sends funds to it. Consequently, only the recipient can compute the private key for this Stealth Address and control the funds.
    \item \textbf{Academic Assessment and Gas Defects:}
          Kovács and Seres (2024) conducted a rigorous academic audit of Umbra in their paper \textit{Anonymity Analysis of the Umbra Stealth Address Scheme on Ethereum}.
          They identified a critical privacy vulnerability—the Gas Problem \cite{kovacs2024}.
          After receiving USDC, the stealth address itself lacks ETH to pay the gas fees required to transfer these USDC. If the recipient transfers ETH from their main wallet to the stealth address, anonymity is instantly compromised (via Gas Payer Linking).
\end{itemize}

\paragraph{Privacy Pools}
Following the sanctions imposed on Tornado Cash, Vitalik Buterin, Jacob Illum et al.
(2024) proposed the concept of “privacy pools” in an attempt to strike a balance between privacy and compliance\cite{buterin2024}.
\begin{itemize}
    \item \textbf{Mechanism Principle:}
          Privacy pools no longer pursue “full-set anonymity,” but instead introduce the concept of Association Sets.
          Users leverage zero-knowledge proofs (zk-SNARKs\footnote{Zero-Knowledge Succinct Non-Interactive Arguments of Knowledge}) to prove:
          \textit{1) I hold a deposit in the pool.}
          \textit{2) My deposit originates from a specific “Proof of Inclusion” set.}
          \textit{3) Alternatively, my deposit origin does not belong to any known “Proof of Exclusion” set of hacked/sanctioned addresses.}
    \item \textbf{SeDe:}
          This solution implements Selective De-anonymization (SeDe), enabling honest users to prove their innocence while severing ties with illicit funds.\cite{sahu2024sede}
    \item \textbf{Anonymous Collection Fragmentation Risk:}
          Academic circles express concern that this mechanism may lead to fragmentation of anonymity sets.
          If each user selects different association sets, the scale (k-anonymity) of each set will significantly decrease, thereby reducing overall privacy protection.
          Furthermore, Buterin et al. (2024) also acknowledge that maintaining association sets introduces centralized or federated trust assumptions (who defines “good” and “bad”?), which could become a new point of entry for censorship \cite{buterin2024}.
\end{itemize}

\section{Privacy Leakage via Collateral, Liquidation, and Position Transparency}

Collateralized stablecoin systems introduce a distinct privacy surface compared with simple token transfers: risk management requires on-chain transparency about positions, yet that same transparency leaks sensitive financial information. In a typical over-collateralized design, users lock collateral into a vault (or equivalent position) and mint stablecoins as debt. The protocol continuously compares the vault’s collateral-to-debt ratio against a liquidation threshold; if the position becomes undercollateralized, it is liquidated via an on-chain auction mechanism. MakerDAO’s Multi-Collateral Dai (MCD) system provides a canonical example of this design: users generate Dai by depositing approved collateral into Maker Vaults, and undercollateralized Vaults are liquidated through automated protocol auctions involving external bidders (``Keepers'') \cite{maker2020mcd}.

\subsection{Threat Model}
The primary adversaries are passive observers and economically-motivated actors who can read and react to public state: blockchain analytics firms, competing market participants, liquidation bots/keepers, MEV searchers, and counterparties. Their capabilities include (i) reading complete on-chain vault state (collateral type, collateral amount, debt, historical adjustments), (ii) monitoring mempools to anticipate liquidations and position management transactions, and (iii) correlating addresses across protocols and exchanges using tagging datasets. Unlike privacy leakage in transfers (which often depends on address reuse), collateralized stablecoin leakage is amplified by the protocol itself publishing risk-relevant state.

\subsection{Leakage mechanisms}
First, \emph{position transparency} makes strategy inference straightforward: a vault’s debt, collateral composition, and safety margin reveal leverage preferences, liquidation tolerance, and timing strategies. Over time, these observations enable profiling (e.g., ``this address increases debt after price rallies'' or ``this institution maintains a narrow buffer''). Second, \emph{liquidation process transparency} creates a targeting surface. In MCD, vaults that cross the liquidation ratio are liquidated via auctions where bidders compete to purchase collateral with Dai, and the auction can evolve into a reverse auction to minimize collateral sold, returning leftovers to the original owner when possible \cite{maker2020mcd}. Because liquidation eligibility is computable from public data and oracle prices, third parties can identify ``weak'' positions and time adversarial actions: monitoring near-threshold vaults, front-running debt repayments, or concentrating bidding behavior around particular victims. Third, \emph{cross-protocol linkability} arises because collateral and stablecoins flow through multiple DeFi protocols. Even if a user attempts to obfuscate transfer history, vault interactions (mint, repay, withdraw) form recognizable interaction patterns that can be clustered with other activities. Finally, side channels such as mempool visibility and RPC metadata can reveal the timing and intent of vault adjustments, enabling correlation with off-chain events.

\subsection{Impact}
These leakages harm both retail and institutional adopters. At the user level, public exposure of \emph{how close to liquidation} a vault is can increase expected losses: adversaries can extract value by racing liquidations, influencing auction dynamics, or exploiting execution ordering (MEV). At the market-structure level, transparency can deter institutions from using decentralized stablecoin minting for treasury management because it reveals balance sheet exposure and risk posture. More broadly, protocol-level transparency creates a privacy--stability tension: the system benefits from verifiable solvency and observable risk, but individual users incur confidentiality costs. This tension can push high-value users toward centralized venues or private settlement, reducing the open stablecoin ecosystem’s liquidity and composability.

\subsection{Mitigations}
A natural direction is to hide \emph{position details} while retaining \emph{verifiable constraints}. One approach is Zero-Knowledge (ZK) proofs that attest to correctness without revealing the underlying values. Zerocash demonstrates how zk-SNARKs can be used to prove validity of complex ledger statements while hiding origin, destination, and amount information \cite{ben2014zerocash}. While stablecoin vault constraints differ from payment constraints, the same principle applies: a user could prove in ZK that (i) their position is well-formed, (ii) the collateralization ratio is above the liquidation threshold, and (iii) any mint/repay operation preserves the protocol’s accounting, without revealing exact collateral or debt. In practice, range proofs are particularly relevant because collateralization checks boil down to inequalities. Bulletproofs provide short, efficient, and aggregatable range proofs with logarithmic proof size, enabling a prover to show a committed value lies in a range without revealing it \cite{bunz2018bulletproofs}. This can support \emph{minimal disclosure} proofs (e.g., proving debt is below a bound, or collateral value exceeds a threshold) that preserve privacy while retaining public verifiability.

Complementary protocol-level mitigations focus on the liquidation pathway. For example, batch auctions, delayed revelation of liquidation targets, and commit--reveal bidding can reduce targeted liquidation and mempool-based sniping, though they must be balanced against price discovery and timely recapitalization. Selective disclosure mechanisms (auditor/view keys) can support compliance and dispute resolution without making position data universally public, but they introduce governance and key-management complexity. Finally, wallet and UX-level measures, such as address hygiene, avoiding address reuse for vault controllers, privacy-aware RPC usage, reduce cross-protocol correlation but cannot fully address leakage rooted in protocol-required transparency.

\subsection{Summary}
In collateralized stablecoin systems, privacy leakage is not merely a byproduct of user behavior; it is structurally induced by on-chain risk management and liquidation transparency. Mitigations should therefore combine cryptographic minimal-disclosure proofs (zk-SNARK-style validity proofs \cite{ben2014zerocash} and efficient range proofs \cite{bunz2018bulletproofs}) with execution-layer and auction-design changes, aiming to preserve both verifiable solvency and user confidentiality.
\section{Cross-layer and Cross-chain Correlation (L2 and Bridges)}

\subsection{Threat Model}
Stablecoin users nowadays frequently move assets across Layer 1 (L1), Layer 2 (L2), and cross-chain bridges for payments, liquidity management, and settlement. These transfers are often routine and repeated, making them observable over time. We consider a malicious user who can monitor public on-chain data across multiple blockchains, including bridge contract events and rollup batch submissions, and correlate these observations across domains.
Beyond raw on-chain records, this user may also exploit coarse-grained external signals, such as transaction timing regularities and commonly followed transfer conventions. When combined, these signals strengthen cross-domain inference. Recent studies show that publicly available bridge and rollup data alone are sufficient to support large-scale linkage analysis across chains and layers, even without privileged access or insider information~\cite{tyan2025,kyan2025}.

\subsection{Attack Mechanisms}
The primary attack mechanism combines \emph{linkable anchors} with heuristic matching. Cross-chain bridges expose structured events (e.g., lock/mint or burn/release) that naturally connect transactions on different chains. When transactions exhibit similar timestamps, token types, and transfer amounts within a limited time window, attackers can match source- and destination-side transfers with high confidence, and further expand these links through address-level clustering and graph propagation~\cite{tyan2025,kyan2025}.
This risk is particularly pronounced for stablecoins, whose transfers often follow standardized denominations and repetitive usage patterns, making amount- and timing-based heuristics more effective. More generally, prior work shows that simple heuristics can become powerful in practice when users reuse addresses or follow predictable transaction routines, amplifying linkability at the entity level~\cite{mharrigan2016}.
In real-world settings, attribution may be further strengthened when linked transaction paths intersect with externally labeled entities (e.g., regulated on/off-ramps), which can act as additional anchoring points for identity inference. Industry analyses of illicit fund flows indicate that cross-chain routes are frequently used to fragment and reroute assets, paradoxically creating more observable linkage points for post hoc tracing~\cite{chainalysis2025}.

\subsection{Impact}
Cross-domain linkability substantially weakens the practical privacy of stablecoin users. While individual transactions may appear innocuous in isolation, aggregated observations across L1, L2, and bridges can reveal sensitive commercial relationships, payment routines, and financial behaviors. Such exposure increases risks of competitive intelligence leakage, targeted surveillance, and user-level security threats.
At the same time, privacy-enhancing mechanisms that aim to provide strong anonymity at the transaction level face significant regulatory and legal uncertainty. Recent enforcement actions and subsequent legal developments surrounding Tornado Cash illustrate how compliance boundaries for privacy technologies can shift rapidly, complicating the deployment of privacy-heavy designs in mainstream stablecoin ecosystems~\cite{reuters2025}.

\subsection{Mitigations}
Mitigation strategies generally follow two complementary directions. One approach is to reduce inherent linkability through design choices, such as batching bridge transactions, introducing controlled temporal delays, or standardizing transfer amounts. These techniques enlarge anonymity sets without modifying the underlying asset model. Limiting metadata exposure at rollup and RPC interfaces can further constrain cross-domain correlation.
A second direction focuses on compliance-friendly privacy mechanisms, which are more suitable for stablecoins intended for large-scale adoption. Approaches based on selective disclosure and verifiable credentials allow users to demonstrate regulatory attributes—such as acceptable risk status or exclusion from sanction lists—without revealing complete transaction histories. This "privacy with accountability" paradigm is increasingly emphasized in both technical research and policy discussions as a proper balance between confidentiality and AML/CFT\footnote{Combating the Financing of Terrorism} requirements~\cite{dduffie2025}.

\section{Compliance-Driven Identity Disclosure}

Within digital asset markets, fiat-pegged stablecoins such as Tether (USDT) and USD Coin (USDC) have increasingly replaced direct fiat currency pairs as the dominant medium of exchange and settlement layer. Recent evidence suggests that stablecoin-denominated trading accounts for approximately 80\% of total cryptocurrency trading volume, underscoring their central role in the digital asset ecosystem \cite{waller2025,theblock2025}. This structural shift positions stablecoins as \emph{de facto} on-ramps and off-ramps for digital assets, thereby amplifying concerns surrounding privacy erosion and compliance-driven control. In this section, we will use USDT as a case study, to examine the divergence between original design assumptions and current operational reality, followed by a discussion of proposed mitigation approaches.

\subsection{Threat Model}
In this section, the main adversary is a \emph{privileged compliance actor}, including stablecoin issuers and regulated intermediaries (VASPs) that can combine on-chain observability with off-chain identity datasets (KYC records) and administrative controls (e.g., freezing or blacklisting). We also consider regulators and analytics firms that rely on these compliance datasets to link transactions to real-world entities.

\subsection{Problem: Privacy, KYC, and Regulatory Drift in USDT}

USDT was originally proposed as a blockchain-based representation of fiat currency, designed to facilitate seamless peer-to-peer transactions without intermediaries \cite{tether2014}. However, over time, Tether Limited has increasingly adopted centralized control mechanisms, including KYC protocols and transaction monitoring systems, ostensibly to comply with evolving regulatory requirements. This shift has led to a divergence between the original vision of USDT as a privacy-preserving digital asset and its current operational reality as a regulated financial instrument.

While USDT transactions are recorded on public blockchains, Tether Limited maintains the ability to freeze or blacklist addresses associated with illicit activities, effectively undermining the pseudonymous nature of blockchain transactions. Furthermore, the implementation of KYC procedures for USDT issuance and redemption introduces additional vectors for privacy leakage, as user identities can be linked to on-chain activities. These compliance measures are driven by regulatory frameworks such as FinCEN's guidance on virtual currencies \cite{finCEN2019}, FATF's risk-based approach to virtual assets \cite{fatf2021}, and OFAC's sanctions compliance guidance \cite{ofac2021}.

Furthermore, despite originally deployed over the Bitcoin blockchain via the Omni Layer protocol, USDT has since expanded to multiple blockchains, including Ethereum, Tron, and Solana. Each of these platforms has distinct privacy characteristics and regulatory compliance mechanisms, further complicating the privacy landscape for USDT users. The multi-chain deployment fragmented liquidity and complicated regulatory oversight, increasing reliance on off-chain intermediaries for compliance enforcement rather than deterministic compliance guarantees \cite{micar2023}.

These factors collectively contribute to a complex privacy environment for USDT users, where the interplay between on-chain transparency and off-chain regulatory compliance creates significant challenges for maintaining user privacy. In the subsequent subsection, we will explore potential mitigation strategies and design trade-offs to address these challenges.

\subsection{Mitigation Approaches and Design Trade-offs}

Regulatory frameworks worldwide mandate KYC and AML compliance for financial service providers, including stablecoin issuers, making complete anonymity incompatible with legal operation in most jurisdictions. The challenge, therefore, is not to eliminate compliance requirements but to design systems that satisfy regulatory obligations while minimizing privacy intrusions. Existing literature proposes several approaches to achieve this balance.

One approach embeds compliance verification mechanisms at the protocol layer of stablecoin systems. Recent work by Duffie et al. \cite{duffie2025} proposes that stablecoin architectures can reconcile strong privacy protections with regulatory obligations (including sanctions screening, KYC, AML, and CFT requirements) by integrating privacy-preserving verification directly into the distributed ledger design, thereby avoiding the current reliance on centralized off-chain intermediaries.

A second approach leverages cryptographic techniques for selective disclosure, allowing users to prove compliance attributes without exposing underlying transaction data or identities. Baldimtsi et al. \cite{baldimtsi2024} systematize over a decade of privacy-preserving payment systems, examining how zero-knowledge proofs and related cryptographic methods enable different privacy guarantees ranging from confidentiality to k-anonymity and sender-receiver unlinkability. However, their analysis identifies a fundamental trilemma: existing systems cannot simultaneously achieve strong anonymity, scalable on-chain data growth, and efficient light-client verification. This limitation suggests that stablecoin designers must make explicit trade-offs between privacy strength, system scalability, and accessibility for resource-constrained users.

A third category employs tiered or hybrid KYC models, as outlined in regulatory frameworks such as FATF's risk-based guidance \cite{fatf2021}. Under this approach, Virtual Asset Service Providers (VASPs) may apply lighter Customer Due Diligence (CDD) procedures for occasional transactions below specified thresholds, such as the USD/EUR 1,000 benchmark in FATF guidance, while requiring full identification and enhanced due diligence for higher-value or higher-risk activities. However, recognizing the unique risk profile of virtual assets, including rapid global settlement and potential for anonymity, some jurisdictions opt to mandate comprehensive KYC regardless of transaction size. This flexibility allows regulatory regimes to calibrate privacy-compliance trade-offs based on jurisdiction-specific risk assessments.

While no single approach fully resolves the fundamental privacy-regulation tension, these mechanisms suggest that future stablecoin architectures may achieve improved privacy without sacrificing regulatory compatibility, provided that design decisions appropriately balance competing policy objectives.

\section{Conclusion}
Stablecoins are now used for payments, settlement, and DeFi minting. This paper shows a simple point: stablecoin privacy does not fail only because users make mistakes. It also fails because the system is built to be transparent.

We discussed four types of privacy leakage. First, transfers are easy to trace on account-based chains. ERC-20 logs and address reuse make linkage cheap. Graph clustering and RPC timing signals can also connect on-chain activity to real-world identities \cite{victor2020,wang2025}. This matters for payroll, merchant settlement, and OTC trades, because repeated payments quickly become a profile. Second, collateralized stablecoins expose positions by design. Vault states, liquidation thresholds, and auctions reveal how users manage leverage and risk. This can attract bots and MEV-driven behavior, and it can raise expected losses near liquidation \cite{maker2020mcd}. Third, moving funds across L1, L2, and bridges does not automatically protect privacy. Bridge events act like anchors. Timing and amount matching can reconnect the path across domains \cite{tyan2025,kyan2025}. Fourth, compliance adds another privacy layer that is specific to stablecoins. Issuers and intermediaries often hold KYC records. They can also freeze or blacklist addresses. So on-chain pseudonyms can become weak in practice under AML and sanctions regimes \cite{finCEN2019,fatf2021,ofac2021,micar2023}.

These risks point to a trade-off that is hard to avoid. Privacy is good for users. Auditability is good for trust and oversight. Risk management needs observability, especially for collateral and liquidation. Regulation also pushes systems toward disclosure. If a design maximizes transparency, privacy collapses. If a design maximizes secrecy, accountability becomes harder. So the real goal is to set boundaries and reduce unnecessary leakage. Systems should hide what is not needed, and prove what must be true. Selective disclosure ideas move in this direction \cite{sahu2024sede,duffie2025}.

Future work can follow three practical directions. First, ZK-based payments and constraint proofs still need better performance and lower costs, so privacy can work at scale without bad UX \cite{ben2014zerocash,bunz2018bulletproofs}. Second, we need compliance-friendly privacy standards, such as interoperable credentials, so users can prove required attributes across wallets and providers without exposing full histories \cite{fatf2021,duffie2025}. Third, cross-chain privacy should be studied together with MEV and execution-layer threats. Bridging creates linkage points, and ordering games amplify harm. Mitigation will likely need joint designs that combine unlinkability, batching, and MEV-aware execution instead of treating them as separate problems.


% References are listed in the file named "references.bib"
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
